Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-06-24 ltx/PLoS/LaTeX_Manuscript.tex
some table edits
some general edits

% Leave date blank
\date{}

@@

   	\caption{{\bf Mexican-hat connectivity kernel, $w(\mathbf{r},\mathbf{r'})$, used in this paper}. The function $w(.)$ is rotationally symmetric (isotropic) about zero, hence a cross-section captures the important aspect of the connectivity kernel's shape. The kernel decays asymptotically to zero.}

@@

   	\caption{{\bf An example of a continuous one-dimensional field decomposed by a finite number of basis functions scaled by the state vector}.} 

@@

\subsection*{From Appendix: Discrete Time Model}\label{Time Discretization} To form the IDE neural field model a time discretisation must be performed. We used a one-step Euler method where equation~\ref{FinalFormContinuous} can be approximated by 
\begin{eqnarray}
	\label{Euler Approximation} \lefteqn{\frac{v\left( \mathbf{r},t+T \right) - v\left( \mathbf{r},t\right)}{T_s} =}\nonumber\\
& -&\zeta v\left( \mathbf{r},t \right) + \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}.\nonumber \\ 
\end{eqnarray}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~\ref{Euler Approximation} we get 
\begin{eqnarray}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) &=& v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right)\nonumber \\
&+& T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}.\nonumber \\ 
\end{eqnarray}
The discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1} v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.

\subsection{From Appendix: Mean and covariance of the state disturbance }\label{ColoredNoise} 
The state disturbance defined in equation~\ref{eq:Wt} is a linear function of $e_t(\mathbf r)$, therefore $\mathbf{e}_t$ is also normally distributed. The expected value of $\mathbf e_t$ is given by 
\begin{eqnarray}
	\mathbf E\left[ \mathbf e_t\right]&=& \mathbf{\Gamma}^{-1}\int_{\Omega}\boldsymbol\phi\left(\mathbf{r}\right)\mathbf E\left[e_t\left(\mathbf{r}\right)\right] d\mathbf{r} \nonumber \\
	&=&\mathbf 0.
\end{eqnarray}
The covariance of $\mathbf{e}_t$ is 
\begin{eqnarray}
	\boldsymbol\Sigma_e =\lefteqn{\mathbf E\left[ \mathbf e_t\mathbf e_t^\top\right]} \nonumber \\ 
&=&\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r})d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}^{\top}(\mathbf{r}') e_t(\mathbf{r}')d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega} \boldsymbol{\phi}(\mathbf{r}) \mathbf E[e_t(\mathbf{r})e_t(\mathbf{r}')]\boldsymbol{\phi}^{\top}(\mathbf{r}')d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
	&=&\mathbf{\Gamma}^{-1}\int_{\Omega}\int_{\Omega}\boldsymbol{\phi}(\mathbf r) \gamma(\mathbf r- \mathbf r' )\boldsymbol{\phi}^{\top}(\mathbf r')d\mathbf r' d\mathbf r\mathbf{\Gamma}^{-\top}.
\end{eqnarray}

\subsection*{From Appendix: Spatial Cut-off Frequency of Field Basis Functions}\label{ap:FrequencyAnalysis}
This Appendix provides the derivation of the field basis function cut-off frequency as a function of the basis function width (equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} in the main text). Applying the \textit{n}-dimensional Fourier transform \cite{Arsac1966} to a \textit{n}-dimensional Gaussian centred at the origin (defined in equation~\ref{eq:BasisFunctionAtOrigin}) yields
\begin{eqnarray}\label{eq:AppendixGaussianFT}
 \lefteqn{\boldsymbol\Phi(\boldsymbol \nu)=\int_{\mathbb{R}^n}{\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2}\mathbf r^\top\mathbf r}\right)\mathrm{exp}\left(-2\pi i\boldsymbol\nu^\top\mathbf r\right)d\mathbf r}} \nonumber \\
&=&\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r \nonumber \\
&\times& \mathrm{exp}\left(-\sigma_{\phi}^2\pi^2\boldsymbol\nu^\top \boldsymbol\nu\right).
\end{eqnarray}
Now evaluating the integral on the RHS of equation~\ref{eq:AppendixGaussianFT} gives 
\begin{eqnarray}\label{eq:IntegralOfGaussian}
\int_{\mathbb{R}^n}\mathrm{exp}\left(-\frac{1}{\sigma_{\phi}^2}\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]^\top\left[\mathbf r +\sigma_{\phi}\pi i \boldsymbol\nu\right]\right)d\mathbf r&& \nonumber \\
=\left(\pi\sigma_{\phi}^2\right)^{\frac{n}{2}}.&&
\end{eqnarray}
Substituting equation~\ref{eq:IntegralOfGaussian} into equation~\ref{eq:AppendixGaussianFT} gives the scaled Gaussian
\begin{equation}\label{eq:FTofGaussian}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)^{\frac{n}{2}}\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation} 
where 
\begin{equation}
	\sigma_{\nu}^2=\frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}

For $3$~dB attenuation ($1/2$ to peak power) at the cut-off frequency, $\boldsymbol\nu_c$, we need to set
\begin{eqnarray}
 \left|\boldsymbol\Phi(\boldsymbol\nu_{c})\right|^2=\frac{1}{2}\left|\boldsymbol\Phi(\mathbf 0)\right|^2,
\end{eqnarray}
since the maximum of Gaussian described in equation~\ref{eq:FTofGaussian} is at the origin. Now solving for $\sigma_{\nu}^2$ we get
\begin{equation}
 \sigma_{\nu}^2=\frac{2\boldsymbol\nu_{c}^\top \boldsymbol\nu_{c}}{\ln 2 }.
\end{equation}

\subsection*{From Appendix: Convolution of Two $n$-Dimensional Gaussians}\label{ap:ConvOfGaussians}
In this appendix we provide derivation for the convolution of Gaussian basis functions. This derivation is used in the analytic calculation of $\boldsymbol\Psi(\mathbf{r})$ and the observation matrix, $\mathbf C$. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}
The derivation proceeds by assuming the Gaussian basis functions are positioned at the origin ($\boldsymbol\mu_i=\boldsymbol\mu_j=\mathbf 0$). The convolution theorem states that the Fourier transform of the convolution of two functions is the product of their Fourier transforms. Using the derivation of the Fourier transform of a \emph{n}-dimensional Gaussian from equation~\ref{eq:FTofGaussian} of \ref{ap:FrequencyAnalysis} we have
\begin{eqnarray}
\mathcal{F}\left(\varphi_i*\varphi_j\right)(\mathbf{r}) &=& \left(\pi\sigma_i^2\sigma_j^2\right)^{\frac{n}{2}}\left(\frac{1}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} \nonumber \\
&\times& (\pi(\sigma_i^2+\sigma_j^2))^{\frac{n}{2}}\mathrm{exp}\left(-(\sigma_i^2+\sigma_j^2)\pi^2\boldsymbol\nu^\top\boldsymbol\nu\right),
\end{eqnarray}
where $\mathcal{F}$ denotes the Fourier transform. Taking the inverse Fourier transform, the convolution is
\begin{equation}\label{eq:ConvoftwoGaussian}
 \left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} \mathbf r^\top\mathbf r}\right).
\end{equation}
In the case where $\boldsymbol\mu_i\neq\boldsymbol\mu_j\neq\mathbf 0$ equation \ref{eq:ConvoftwoGaussian} becomes
\begin{equation}\label{eq:ConvoftwoGaussianNonzeroMean}
 \left(\varphi_i*\varphi_j\right)(\mathbf{r}) = \left( \frac{\pi\sigma_i^2\sigma_j^2} {\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}} \mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\mathbf r-\boldsymbol\mu)^\top(\mathbf r-\boldsymbol\mu)}\right).
\end{equation}
where $\boldsymbol\mu=\boldsymbol\mu_i+\boldsymbol\mu_j$. 
The analytic definition of the observation matrix, $C$, and the matrix $\boldsymbol\Psi(\mathbf{r})$ are formulated using this derivation.

\section*{From Appendix: Inner Product of Two \emph{n}-Dimensional Gaussians}\label{ap:InnerProdOfGaussians}
The inner product of the two \emph{n}-dimensional Gaussians (from equations~\ref{eq:n_dimensional_Gaussian1} and~\ref{eq:n_dimensional_Gaussian2}) is defined as
\begin{equation}\label{eq:InofGaussianDefinition}
\varphi_i\otimes\varphi_j=\int_{\mathbb{R}^n}\varphi_i(\mathbf r)\varphi_j(\mathbf r)d\mathbf{r}.
\end{equation}
A Gaussian basis function is symmetric with respect to its centre, therefore
\begin{equation}\label{eq:GaussianSymmetry}
 \varphi_i(\boldsymbol \mu_i+\mathbf r)= \varphi_i(\boldsymbol \mu_i-\mathbf r).
\end{equation}
Letting $\mathbf{r}= \mathbf{r}-\boldsymbol{\mu}_i$ equation~\ref{eq:GaussianSymmetry} becomes
\begin{equation}\label{eq:ConsequenceofGaussianSymmetry}
 \varphi_i(\mathbf r)= \varphi_i(2\boldsymbol \mu_i-\mathbf r).
\end{equation}
Now substituting equation~\ref{eq:ConsequenceofGaussianSymmetry} into equation~\ref{eq:InofGaussianDefinition} gives 
\begin{eqnarray}
\varphi_i\otimes\varphi_j&=&\int_{\mathbb{R}^n}\varphi_i(\mathbf r)\varphi_j(\mathbf r)d\mathbf r\\
&=&\int_{\mathbb{R}^n}\varphi_i(2\boldsymbol \mu_i-\mathbf r)\varphi_j(\mathbf r)d\mathbf r\\
&=&(\varphi_i*\varphi_j)(2\boldsymbol \mu_i).
\end{eqnarray}
Evaluating equation~\ref{eq:ConvoftwoGaussianNonzeroMean} at $\mathbf r=2\boldsymbol\mu_i $ gives
\begin{equation}
 \varphi_i\otimes\varphi_j=\left(\frac{\pi\sigma_i^2\sigma_j^2}{\sigma_i^2+\sigma_j^2}\right)^{\frac{n}{2}}\mathrm{exp}\left({-\frac{1}{\sigma_i^2+\sigma_j^2} (\boldsymbol\mu_i-\boldsymbol\mu_j)^\top(\boldsymbol\mu_i-\boldsymbol\mu_j)}\right).
\end{equation}

@@

    	\caption{{\bf \textbf{(A)} Example of the neural feld}. (\textbf{B}) Example of data generated by the full neural field model from the first 5 sensors. (\textbf{C}) Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 

@@

\begin{table}[!ht]
\caption{
\bf{Parameters}}
\begin{tabular}{|c|l|c|c|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Value} & \textbf{Units}\\
	\hline
	\multicolumn{4}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Delta$ & spatial discretisation step & 0.5 & mm \\
	$T_s$ & time step & $0.001$ & s\\
	$T$ & number of time steps used in estimation & 0.4 & s\\
	\hline 
\multicolumn{4}{|c|}{\emph{Model}}\\
	\hline
	$\tau$ & synaptic time constant & 0.01 & s$^{-1}$ \\
	$f_{max}$ & maximal firing rate & 10 & spikes.s$^{-1}$\\
	$\varsigma$ & slope of sigmoidal activation function & 0.8 & mV$^{-1}$\\
	$v_0$ & firing threshold & 2 & mV \\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\end{array}
	\right]^{\top}$ & mV.spike$^{-1}$\\
	$\sigma_{\psi_{1}}, \sigma_{\psi_{2}}, \sigma_{\psi_{3}}$ & connectivity kernel width parameters & 1.8, 2.4, 6 & mm\\
	$\Delta_y$ & distance between adjacent sensors & 1.5 & mm\\
	$\sigma_{m}$ & sensor width parameter & 0.9 & mm \\
	$\Sigma_{\varepsilon}$ & observation noise variance & $0.1 \times I_{N}$ & mm$^2$ \\
	$\sigma_{\gamma}$& disturbance spatial covariance width & 1.3 & mm\\
	$\sigma_{\gamma,t}$ & disturbance temporal variance & 0.1 & mV$^2$ \\
	\hline 
	\multicolumn{4}{|c|}{\emph{Reduced Model}}\\
	\hline
	$\Delta_{\phi}$ & Distance between field basis functions & 2.5 & mm \\
	$\sigma^2_{\phi}$ & Spatial variance of Gaussian in spatial domain & 2.5 & mm$^2$\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Frequency Analysis}}\\
	\hline
	$\rho_y$ & Sensor over-sampling parameter & 1 & n.a.\\
	$\rho_\phi$ & Field basis function over-sampling parameter & 1 & n.a.\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Estimation}}\\
	\hline
	$\alpha$& Determines the spread of sigma points & 0.001 & n.a.\\
	$\beta$& Prior knowledge of the sigma points distribution & 2 & n.a.\\
	$\kappa$& Scaling parameter & $3-L$ & n.a.\\
	\hline
\end{tabular}
\begin{flushleft}Parameters for the neural field model, the model reduction, the frequency analysis and the estimation procedure.
\end{flushleft}
\label{tab:Parameters}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Notation}}
\begin{tabular}{|c|l|c|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain &\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y(\mathbf{r}_n,t)$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$g(\mathbf{r},t)$ & weighted firing rate function & spikes.s$^{-1}$\\
	$f(\mathbf{r},t)$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\gamma(\mathbf r-\mathbf r')$ & mV\\
	$\varepsilon(\mathbf{r}_n,t)$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n,\mathbf{r}')$ & sensor kernel where $n$ is sensor index $n=1,..,N$ (width $\sigma_m^2$) & n.a. \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\zeta$ & inverse synaptic time constant & s$^{-1}$\\
	$\xi$ & time constant parameter & n.a.\\
	$\eta(t)$ & Heaviside step function & n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	$D$ & temporal differential operator & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\boldsymbol\phi(\mathbf{r})$ & vector of $L$ Gaussian basis functions & n.a.\\
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$ & vector of connectivity kernel basis functions & n.a.\\
   	$\Gamma$ & inner product of field basis functions & n.a.\\
   	$q()$ & parameterised nonlinear state function (used in LS estimator) & n.a.\\
   	$Q()$ & state transition function &\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
    $V(\nu)$ & Fourier transform of the neural field & Hz\\
	$\nu$ & Spatial frequency & Hz \\
	$\nu_c$ & Field spatial cut-off frequency & Hz\\
	$\sigma_{\nu}^2$ & Variance of Fourier transformed Gaussian basis function & Hz$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\mathcal{X}$ & matrix of sigma vectors & mV\\
	$W$ & matrix of sigma vector weights (scaling parameter $\lambda$) & n.a.\\
   	$\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^f$ & forward prior and posterior state estimates & mV\\
   	$\hat{\mathbf{x}}_t^{b-}$, $\hat{\mathbf{x}}_t^{b}$ & backward prior and posterior state estimates & mV\\
 	$P^{f-}_t$, $P^f_t$  & forward prior and posterior covariance matrices & mV$^2$\\
   	$P^{b-}_t$, $P^b_t$ & backward prior and posterior covariance matrices & mV$^2$\\
	$M_t$& cross-covariance matrix & mV$^2$\\
	$\mathcal K_{t} $ & filter gain & n.a.\\
	$S_t$ & smoother gain & n.a.\\	
	\hline
\end{tabular}
\begin{flushleft}The first column of the table gives the symbols for all the notation used the paper. The second column provides a brief description of the quantity, and the third column provide the relevant SI unit where appropriate.
\end{flushleft}
\label{tab:label}
\end{table}

>>>>

% Leave date blank
% \date{}

@@

   	\caption{{\bf Mexican-hat connectivity kernel, $w(\mathbf{r},\mathbf{r'})$}. The function $w(.)$ is rotationally symmetric (isotropic) about zero, hence a cross-section captures the important aspect of the connectivity kernel's shape. The kernel decays asymptotically to zero.}

@@

   	\caption{{\bf Example of a field decomposition.} An example of a continuous one-dimensional field decomposed by a finite number of basis functions scaled by the state vector.} 

@@

@@

    	\caption{{\bf \textbf{(A)} Example of the neural field}. (\textbf{B}) Example of data generated by the full neural field model from the first 5 sensors. (\textbf{C}) Power spectral density of the data generated by an observation showing the typical $1/f^2$ characteristic of intracranial EEG.} 

@@

\begin{table}[!ht]
\caption{
\bf{Parameters}}
\begin{tabular}{|c|l|c|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Value} & \textbf{Units}\\
	\hline
	\multicolumn{4}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Delta$ & spatial discretisation step & 0.5 & mm \\
	$T_s$ & time step & $0.001$ & s\\
	$T$ & number of time steps used in estimation & 0.4 & s\\
	\hline 
\multicolumn{4}{|c|}{\emph{Model}}\\
	\hline
	$\tau$ & synaptic time constant & 0.01 & s$^{-1}$ \\
	$f_{max}$ & maximal firing rate & 10 & spikes.s$^{-1}$\\
	$\varsigma$ & slope of sigmoidal activation function & 0.8 & mV$^{-1}$\\
	$v_0$ & firing threshold & 2 & mV \\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters & $\left[\begin{array}{ccc}
	10 &-8 &0.5
	\end{array}
	\right]^{\top}$ & mV.spike$^{-1}$\\
	$\sigma_{\psi_{1}}, \sigma_{\psi_{2}}, \sigma_{\psi_{3}}$ & connectivity kernel width parameters & 1.8, 2.4, 6 & mm\\
	$\Delta_y$ & distance between adjacent sensors & 1.5 & mm\\
	$\sigma_{m}$ & sensor width parameter & 0.9 & mm \\
	$\Sigma_{\varepsilon}$ & observation noise variance & $0.1 \times I_{N}$ & mm$^2$ \\
	$\sigma_{\gamma}$& disturbance spatial covariance width & 1.3 & mm\\
	$\sigma_{\gamma,t}$ & disturbance temporal variance & 0.1 & mV$^2$ \\
	\hline 
	\multicolumn{4}{|c|}{\emph{Reduced Model}}\\
	\hline
	$\Delta_{\phi}$ & Distance between field basis functions & 2.5 & mm \\
	$\sigma^2_{\phi}$ & Spatial variance of Gaussian in spatial domain & 2.5 & mm$^2$\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Frequency Analysis}}\\
	\hline
	$\rho_y$ & Sensor over-sampling parameter & 1 & n.a.\\
	$\rho_\phi$ & Field basis function over-sampling parameter & 1 & n.a.\\
	\hline 
	\multicolumn{4}{|c|}{\emph{Estimation}}\\
	\hline
	$\alpha$& Determines the spread of sigma points & 0.001 & n.a.\\
	$\beta$& Prior knowledge of the sigma points distribution & 2 & n.a.\\
	$\kappa$& Scaling parameter & $3-L$ & n.a.\\
	\hline
\end{tabular}
\begin{flushleft}Parameters for the neural field model, the model reduction, the frequency analysis and the estimation procedure.
\end{flushleft}
\label{tab:Parameters}
\end{table}

\begin{table}[!ht]
\caption{
\bf{Notation}}
\begin{tabular}{|c|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y(\mathbf{r}_n,t)$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$g(\mathbf{r},t)$ & weighted firing rate function & spikes.s$^{-1}$\\
	$f(\mathbf{r},t)$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\gamma(\mathbf r-\mathbf r')$ & mV\\
	$\varepsilon(\mathbf{r}_n,t)$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n,\mathbf{r}')$ & sensor kernel where $n$ is sensor index $n=1,..,N$ (width $\sigma_m^2$) & n.a. \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\zeta$ & inverse synaptic time constant & s$^{-1}$\\
	$\xi$ & time constant parameter & n.a.\\
	$\eta(t)$ & Heaviside step function & n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	$D$ & temporal differential operator & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\boldsymbol\phi(\mathbf{r})$ & vector of $L$ Gaussian basis functions & n.a.\\
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$ & vector of connectivity kernel basis functions & n.a.\\
   	$\Gamma$ & inner product of field basis functions & n.a.\\
   	$q()$ & parameterised nonlinear state function (used in LS estimator) & n.a.\\
   	$Q()$ & state transition function &\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
    $V(\nu)$ & Fourier transform of the neural field & Hz\\
	$\nu$ & Spatial frequency & Hz \\
	$\nu_c$ & Field spatial cut-off frequency & Hz\\
	$\sigma_{\nu}^2$ & Variance of Fourier transformed Gaussian basis function & Hz$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\mathcal{X}$ & matrix of sigma vectors & mV\\
	$W$ & matrix of sigma vector weights (scaling parameter $\lambda$) & n.a.\\
   	$\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^f$ & forward prior and posterior state estimates & mV\\
   	$\hat{\mathbf{x}}_t^{b-}$, $\hat{\mathbf{x}}_t^{b}$ & backward prior and posterior state estimates & mV\\
 	$P^{f-}_t$, $P^f_t$  & forward prior and posterior covariance matrices & mV$^2$\\
   	$P^{b-}_t$, $P^b_t$ & backward prior and posterior covariance matrices & mV$^2$\\
	$M_t$& cross-covariance matrix & mV$^2$\\
	$\mathcal K_{t} $ & filter gain & n.a.\\
	$S_t$ & smoother gain & n.a.\\	
	\hline
\end{tabular}
\begin{flushleft}The first column of the table gives the symbols for all the notation used the paper. The second column provides a brief description of the quantity, and the third column provide the relevant SI unit where appropriate.
\end{flushleft}
\label{tab:label}
\end{table}

\renewcommand{\arraystretch}{1.8}
\begin{table}[!ht]
\caption{
\bf{Steps in the Algorithm for the Unscented RTS Smoother}}
\begin{tabular}{|c|}\hline
\multicolumn{1}{|p{16cm}|}{\textbf{1.} Forward initialisation} \\ 
$\hat{\mathbf x}_0, \mathbf P_0$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{2.} Forward iteration: for $t \in \left\lbrace 0,\cdots, T\right\rbrace $, calculate the sigma points $\mathcal X_{i,t}^f$ using equations \ref{eq:sigmapoints1}-\ref{eq:sigmapoints3} and propagate through equation~\ref{eq:QmatrixForSigmapoints}} \\
$\mathcal X_{i,t+1}^{f-}=Q(\mathcal X_{i,t}^f) \quad i=0, \dots, 2L$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix} \\
$\hat{\mathbf x}_{t+1}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{f-} $ \\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e$ \\ 
\multicolumn{1}{|p{16cm}|}{Compute the filter gain, the filtered state and the filtered covariance matrix using the standard Kalman filter update equations} \\
$\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_{\varepsilon})^{-1}$\\
$\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-}) $\\
$\mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}$\\ 
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{3.} Backward initialisation}\\
$\mathbf P_T^b= \mathbf P_T^f, \quad \hat{\mathbf x}^b_T= \hat{\mathbf x}^f_T$ \\
\hline
\multicolumn{1}{|p{16cm}|}{\textbf{4.} Backward iteration: for $t \in \left\lbrace T-1, \cdots, 0 \right\rbrace $ calculate the sigma points $\mathcal X_{i,t}^b$ and propagate them through equation \ref{eq:QmatrixForSigmapoints}}\\
$\mathcal X_{i,t+1}^{b-}=Q(\mathcal X_{i,t}^b) \quad i=0, \dots, 2L$\\
\multicolumn{1}{|p{16cm}|}{Calculate the predicted state and the predicted covariance matrix}\\
$\hat{\mathbf x}_{t+1}^{b-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1}^{b-}$\\
$\mathbf P_{t +1}^{f-}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e $\\
$\mathbf M_{t +1}=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t}^{b-}-\hat{\mathbf x}_{t}^{f})(\mathcal X_{i,t+1}^{b-}-\hat{\mathbf x}_{t+1}^{b-})^\top$\\
\multicolumn{1}{|p{16cm}|}{Compute the smoother gain, the smoothed state and the smoothed covariance matrix}\\
$\mathbf S_t=\mathbf M_{t +1}\left[ \mathbf P_{t +1}^{b-}\right] ^{-1} $\\
$\hat{\mathbf x}_t^b=\hat{\mathbf x}_t^f+\mathbf S_t\left[\hat{\mathbf x}_{t+1}^{b}-\hat{\mathbf x}_{t+1}^{b-}\right]$\\
$\mathbf P_{t}^{b}=\mathbf P_{t}^{f}+\mathbf S_t\left[\mathbf P_{t+1}^{b}-\mathbf P_{t+1}^{b-} \right]\mathbf S_t^\top $\\
\hline
\end{tabular}
\begin{flushleft}This table shows the steps in the unscented Rauch-Tung-Striebel smoother algorithm. The steps are iterated 11 times for our state estimation procedure. The least squares algorithm is run after each iteration to update the parameter estimates.
\end{flushleft}
\label{tab:UKFAlgorithm}
\end{table}

\renewcommand{\arraystretch}{1}
\begin{table}[!ht]
\caption{
\bf{Estimation Results}}
\begin{tabular}{|c|ccc|}	
\hline\hline 
Parameter & Actual Value & Mean Estimate $\pm$ Std. Dev. & Bias (\%) \\
\hline\hline
$\xi$ & $0.9$ & $0.925\pm 0.004$ & $2.778$ \\
$\theta_0$ & $10$ & $10.228\pm2.794$ & $2.28$\\
$\theta_1$ & $-8$ & $-8.495\pm1.913$ & $-6.187$\\
$\theta_2$ & $0.5$ & $0.545\pm0.077$ & $9$\\
\hline
\end{tabular}
\begin{flushleft}Actual parameter values compared to mean parameter estimates.
\end{flushleft}
\label{tab:Parameters estimates}
\end{table}
